\chapter{Introduction}
\label{chap:intro}

\section{Motivation}

Railway transportation remains relevant even after its long history due to its efficiency, safety, and~sustainability. However, the~most limiting factor has always been the~use of~the~required infrastructure. As such, the~critical problem of~utilizing the~railway infrastructure effectively has always been scheduling. The complexity of~this problem is continuously increasing with the~size and~intricacy of~the~railway system as well as the~demand for its use. 

Traditionally, scheduling was done manually by humans; only with the~advancements of~optimization algorithms and~computational power have humans been aided by computers. This still produces a static schedule that might be unusable if a disruption occurs. This is the~train dispatching problem, where the~decisions are made in~real-time, managing delays and~conflicts as they happen. Real-time dispatching is often still only done by humans and~provides a~significant challenge for~optimization algorithms as they need to find an acceptable solution in~a matter of~minutes or~even seconds.

\section{Goals}
The main goal of~this thesis is to develop a heuristic that is capable of~train dispatching in~real-time with interpretable results; a human should be able to understand the~structure of~the~result to verify its validity.

The heuristic is evaluated on~the~\emph{DISPLIB} train dispatching benchmark library \cite{kloster2025displiblibrarytraindispatching} consisting of~train dispatching problem (TDP) instances from real-world data. This library of~instances was introduced for~an optimization competition.

A significant part of~this thesis is focused on~the~aggregation and the~structural analysis of~the~TDP as presented in~DISPLIB. This is the~main innovation which allows us to use a top-down approach, reducing the~problem to selecting the~routes top-down and~then scheduling the~trains with a more concise model, as opposed to the~bottom-up approach, building up the~solutions by sequentially routing and~resolving conflicts, which is used by the~winning solutions in~the~DISPLIB competition.

\section{Thesis structure}
\begin{minipage}{\textwidth}
This thesis is structured in~the~following chapters with their contents:

\begin{enumerate}
	\item \hyperref[chap:intro]{\emph{Introduction:}} Outlines the~motivation, goals, and~structure of~this thesis.
	%\hyperref[chap:lit_rev]
	\item \hyperref[chap:lit_rev]{\emph{Literature review:}} Review of~the~pertinent literature separated into three parts, literature connected to~DISPLIB, the~most relevant resources inspiring the~heuristic or~providing key theoretical ideas, and~an overview of~the~selected state-of-the-art in~scheduling optimization.
	\item \hyperref[chap:theor]{\emph{Theoretical background:}} Basic theory necessary for~the~description of~the~problem. The~main focus is on~graph topology and~the~use of~alternative and~weighted graphs for~scheduling models.
	\item \hyperref[chap:method]{\emph{Methodology:}} Detailed description of~the~train dispatching problem, aggregation, routing, and~conflict resolution.
	\item \hyperref[chap:results]{\emph{Results:}} Presentation of~the~performance of~the~procedures described in~methodology on~the~DISPLIB instances.
	\item \hyperref[chap:discuss]{\emph{Discussion:}} Discussion of the ideas and procedures presented in the methodology chapter and discussion of~the~results.
	\item \hyperref[chap:concl]{\emph{Conclusion:}} Highlights of~the~main ideas presented in~this thesis and overall evaluation, future work considerations.
\end{enumerate}
\end{minipage}


