\chapter{Discussion}
\label{chap:discuss}

\section{Aggregation}
From tables \ref{tab:preprocess} and~\ref{tab:preprocess_opt}, we can see that the~aggregation of~the~initial data is successful. For all instance groups, except the~\emph{swi}, the~operation to~level ratio ranges from 1.46 to~2.84; this should provide a~significant reduction in the~number of~necessary vertices when creating the~event graph, or other similar structure where we can leverage the~parallel events. In turn, the~\emph{swi} group has an operation to~route ratio of~20.4. This means that the~operations in these instances cannot be aggregated into parallel events, but instead have long stretching routes of~many operations, that are aggregated into a~single routing decision.

The most significant is the~ORR per chunk ratio. While the~\emph{nor} and~\emph{wab} instances had missing or negligible reductions, due to~the~fact that they assign each operation its separate resource, the~3.25 to~3.52 reduction ratios for \emph{smi} and~\emph{swi} instance groups should provide a~significant reduction in the~number of~conflicts that need to~be resolved when scheduling. Because most combinatorial problems are considered NP-hard \cite{garey1979computers}, reducing the~number of~required decisions by~a factor of~over three should result in much faster solution of~these problems.

From table \ref{tab:time_req} we can see that the~time to~do these aggregation steps is negligible, even compared to~the~process of~loading the~instance data itself, thus it is always recommended to~do them before doing any sort of~further solving or analysis.

The only questionable decision is regarding the~use of~parallel merges. While they may further reduce the~number of~necessary chunks, they also need more bookkeeping to~track which parallel route we took. For a~solution that dynamically updates the~routing decision, it might be simpler to~not merge the~parallel chunks in favor of~only increasing the~number of~chunks by~a smaller percentage.

\section{Link graph}
The derivation of~conditions for the~linked conflicts in section \ref{sec:find_links} from the~alternative graph formulation in \ref{sec:conf_links} via the~use of~the~link graph is the~main theoretical innovation presented in this thesis. We believe that it could be used to~analyze and~solve not only the~train dispatching problem, but also any problem that uses resources in a~blocking manner, like the~blocking job-shop problem (BJSP).

\begin{minipage}{\textwidth}
We successfully use the~link graph for both routing and~conflict resolution, showing its potential as~a useful tool. It is important to~note that prebuilding the~entire possible graph ahead of~time to~allow for continuous updates using operation bitmaps is relatively fast, as~you can see from table \ref{tab:time_req}, at~most 8 times the~amount of~time to~load the~instance, usually faster.	
\end{minipage}

From the~routing time you can see that querying the~link graph for conflict chains and~updating it is also relatively fast, certainly faster than augmenting and~searching the~entire alternative graph for inconsistent selections would be.

\section{Routing}
The top-down routing decisions based on opposite conflict chains have proven to~be quite reasonable, the~improvement in solution quality can be clearly seen from table \ref{tab:solve}. While it did increase the~time to~the~first conflict free solution for some of~the~groups, it also enabled us to~solve the~\emph{nor1\_full} and~\emph{nor4} instances, which all timed out for the~random routing assignment. The routing decisions are clearly motivated by~the~structure of~the~problem and~its safety effect of~preventing trains from going in the~opposite direction to~share resources.

There is, however, one main drawback compared to~the~bottom-up approaches of~the~winning DISPLIB solutions, we do not refine the~routes during the~optimization. This causes us to~be stuck in suboptimal routes, meaning that we only improve a~current feasible solution by~tightening the~MIPGap parameter of~the~MILP solver. It is, however, unclear how to~combine our top-down routing approach with refinement, as~the~bottom-up solutions just use the~best current available route without any considerations to~the~overall structure of~the~problem, treating it as~a black-box.

\section{Conflict resolution}
The path-and-cycle formulation has proven to~be quite effective for the~small and~mid-sized instances, getting a~conflict free solution within seconds or even fractions of~a~second. This seems to~be the~main advantage over the~big-M formulation, where it takes them tens or hundreds of~seconds to~arrive at~a~solution \cite{Dubach2025handling}, showing a~significant speed-up, albeit at~the~cost of~suboptimal solutions for some instances.

This is caused by the~fact that they need to create a~time variable for each level at~least, plus two time variables for each chunk, and the~corresponding temporal feasibility constraints. On~top of that, they also need to use many simplex iterations to get a~feasible starting schedule for these variables. Our formulation uses the~event graph to generate the~level times and chunk intervals, thus our initial MILP model is much smaller, only having the~objective delay variables.

\newpage
The failure to~solve the~\emph{nor5} and~\emph{wab} instance group corresponds to~the~fact, that for these instances the~resource chunk aggregation is not applicable, thus increasing the~number of~required conflicts to~be resolved. It should be noted, that the heuristic in \textcite{Dubach2025handling} also struggled with the~same set of~instances, \emph{nor5} and~\emph{wab}, but not with the~3 larger instances in \emph{swi} as~ours did.

While the~gap to~the~best known solutions might seem high, it is important to~note that these are solutions with unlimited time and~computation resources and~thus some gap is expected. The only other point of comparison is that only 8 teams in the~competition provided at least one solution in more than half of the~instance groups. Out of ten total groups, we fully solved five and solved the~majority of instances in two more.

The main problem for the~path-and-cycle formulation is the~bloat of~the~MILP master problem, where we are adding even thousands of~cuts, which in turn slows down the~solution significantly, even when using the~lazy constraint principle. The biggest possible improvement for the~speed of~this optimization problem would be some sensible cut cleanup procedure that would remove the~redundant cuts that are no longer necessary. We have made an effort to~implement such a~cleanup based on~the~removal of~non-binding or dominated cuts, but it hindered the~performance of~the~solver as~these updates were more time-consuming than just simply letting Gurobi lazy addition to~deal with them.

